# **Analyzing Data**

Once your data is cleaned up then it is ready to investigate the data in order to find your insights and how significant they are.

## **Statistical Tests**

### **T-Test**

A t-test is a good way to see if there is a difference between two groups and "quantifying" the difference of the groups. [Learn more](https://towardsdatascience.com/the-statistical-analysis-t-test-explained-for-beginners-and-experts-fd0e358bbb62) about the t-test.

Lets say I wanted to know if there was a difference between the US and Netherlands when it comes to life expectancy. My next steps would be to create a **null hypothesis** and the **alternative hypothesis**. For this type of question the **null** would be that the averages of both groups are equal, the alternative is that they are not equal.

> Null hypothesis: `average life expectancy of US = average life expectanvy of the Netherlands`
>
> Alternative hypothesis: `average life expectancy of US ≠ average life expectanvy of the Netherlands`

Next I want to select my **critical value** which gives "boundaries" to the acceptance of the test. A common critical value that is used is **0.05**, and this means that there is a 5% chance if done 100 times, that I will reject the null hypothesis and accept the alternative. Critical values such as 0.01 is even more accurate as there is a 1% chance of rejecting the null, but a 99% chance of retaining the null, so choosing a critical value can be a balancing act.

> Critical Value: `𝛼 = 0.05`

Now to run the t-test to see the results.

```{r warning=FALSE, message=FALSE}
# load packages: gapminder for data + tidyverse meta package
library(gapminder)
library(tidyverse)

# filter data for life exp and countries: US and Netherlands
df1 <- gapminder %>%
  select(country, lifeExp) %>%
  filter(country == "United States" | country == "Netherlands") 
  
 t.test(data = df1, lifeExp ~ country)
```

The results of our t-test show us that the Netherlands has a higher average life expectancy than the US, however with a `p-value of .086` I will retain my null hypothesis as there is not a statistical significance in the difference of the means for each of the countries.

What if I look at the life expectancy between the US and South Africa with the same parameters as before?

```{r warning=FALSE, message=FALSE}
# filter data for life exp and countries: US and South Africa
df2 <- gapminder %>%
  select(country, lifeExp) %>%
  filter(country == "United States" | country == "South Africa") 
  
 t.test(data = df2, lifeExp ~ country)
```

As you can see there is a much larger difference in the average life expectancy between the countries, and with a `p-value` of `3.392e-09`. This p-value is in scientific notation as it is such a small decimal, in decimal form it would be `0.000000003392` which is much smaller than .05 meaning this difference in life expectancy is statistically significant. In this case I would reject my null hypothesis and accept the alternative hypothesis that the average life expectancy of the United States is not equal to the average life expectancy of South Africa.

If I where reporting this insight, I could say that the life expectancy in the United States is significantly higher than that of South Africa... and instead of giving the p-value in decimal format or scientific notation, I could say `p < .05`.
